On modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-08-24 docs/mvfits_paper/mvf.tex
misc updates

The forward process is to end when the model accuracy (or changes thereof) pass a specified threshold.

@@

This rough algorithmic picture encapsulated in Alg. \ref{alg:pgreedy}.

@@

\par In this setting, the basis symbols required by Alg. (\ref{alg:pgreedy}) are the multinomial terms in \eqn{eq:mvt}.

@@

        \phi_k(\vec{x}_j)\, g , & \quad \;\; \;\text{for }R+1\leq k\leq R+V

@@

where, here, $\hat{P}$ is the pseudo-inverse of the matrix whose elements are $\psi_k(\vec{x}_j)$.

@@

In this contemplation, it may be that shown that $n$ may be entirely regelated to $\hat{P}$.

@@

It is in this sense that \eqn{eq:pinv2} is insufficient to generally solve for $\vec{\alpha}$.

@@

That is, to reduce the impact of noise on $\hat{P}$, we are free to calculate it using model evaluations of $g$ rather than the original training data.

@@

To this end, if we define $g^{(0)}=g$ (i.e. $g$ is the training data) then, with $\hat{P} = \hat{P}( \vec{g}^{\,(n)} )$, then \eqn{eq:pinv2} generalizes to

@@

In practice, one solves \eqn{eq:pinv3} for $\vec{\alpha}^{(n)}$, and then uses the related $a_r^{(n)}$ and $b_v^{(n)}$ to calculate $g^{(n+1)}$ via

@@

To broach these complications, we again use a greedy algorithm with degree tempering.

@@

Put another way, rather than two simultaneous greedy optimizations over $R+1$ and $V$ symbols (with $(R+1)V$ iterations), a single greedy process over $V+R+1$ symbols is performed.

@@

Both \gmvp{} and \gmvr{} are publicly available on Github through the \texttt{positive} repository \cite{lionel_london_2018_1402516}.

@@

Second, we apply \gmvr{} to the modeling of spin -2 spherical-spheroidal harmonic mixing coefficients.

@@

For relevance of presentation to physics examples in subsequent sections, we will restrict ourselves to a case where numerical noise is low, and the negative greedy step does not alter the output of Alg. \ref{alg:gmvr}).

@@

\par To generate the test data, \eqn{eq:gmvrtoy1} is evaluated with 25 points in $x_0$ and $x_1$, where each is between -3 and 3.

@@

\par In this rudimentary example case, \gmvr{} correctly recovers the functional form of the input data. But, in general, \gmvr{} and related techniques, having no knowledge of the underlying noise distribution, will attempt to model minor correlations and offsets within the training data's noise.

@@

However, we have demonstrated the utility of \gmvr{} in a relatively ideal usage case where the underlying function is rational, and the training data is inly weakly contaminated with noise.

@@

\par In the following sections, we consider still ideal cases, where the functional form of the sample data is not known to be explicitly polynomial or rational, but the amount of noise within the training data is negligible.

@@

\subsection{Modeling \qnm{} frequencies using \gmvp{}}

@@

\par \fig{fig:qnm}'s top left and right panels' upper axis demonstrate the effect of mapping $\jf$ onto $\kappa$.

@@

In particular, it is shown that the two solution branches ($\jf>0$ and $\jf<0$) naturally form a single family of solutions when accounting for the sign of the \bh{}'s oriented spin \cite{Husa:2007hp}.

@@

Concurrently, the use of $\kappa$ as a domain variable has the desirable effect of making each $\cw\lmn$ and $\tau\lmn$ approximately quarctic.

@@

\subsection{Modeling spherical-spheroidal inner-products using \gmvr{}}

@@

For efficiency of presentation, each $\sigma_{\LMlmn}$ is plotted against its real and imaginary part.

@@

In particular, the \gmvp{} algorithm performs multivariate polynomial modeling of real and complex valued scalar functions.

@@

@@

\par Treating a toy problem with \gmvr{} demonstrates \gmvr{}'s ability to faithfully recover underlying model parameters for a plausible dataset.

@@

\par Both \gmvp{} and \gmvr{} may be used to automatically determine the functional form and model for a given dataset that is expected to be respectively multinomial or rational.

@@

These models are found to have residual errors well below $1\%$ when $| \jf | \leq \CwFitCalibrationRegion$.

>>>>

The forward process is to end when the model accuracy, or changes thereof, pass a specified threshold.

@@

This rough algorithmic picture is encapsulated by Alg. \ref{alg:pgreedy}.

@@

\par Here, the basis symbols required by Alg. (\ref{alg:pgreedy}) are the multinomial terms in \eqn{eq:mvt}.

@@

        \phi_k(\vec{x}_j)\, g_j , & \quad \;\; \;\text{for }R+1\leq k\leq R+V

@@

where, $\hat{P}$ is the pseudo-inverse of the matrix whose elements are $\psi_k(\vec{x}_j)$, $\vec\alpha = ( z_0, z_1, ... z_{R+V-1},z_{R+V} )$, and $\vec{g} = ( g_0, g_1, ... g_{R+V-1},g_{R+V} )$.

@@

In this, it may be that shown that $n$ may be entirely regelated to $\hat{P}$.

@@

It is in this sense that \eqn{eq:pinv2} is insufficient to generally solve for $\vec{\alpha}$, as $\hat{P}$ may be adversely affected by noise.

@@

That is, to reduce the impact of noise on $\hat{P}$, we are free to calculate it using model evaluations of $g$ rather than the original (noisy) training data.

@@

To this end, if we define $g^{(0)}=g$ (i.e. $g$ is the training data), with $\hat{P} = \hat{P}( \vec{g}^{\,(n)} )$, then \eqn{eq:pinv2} generalizes to

@@

In practice, one solves \eqn{eq:pinv3} for $\vec{\alpha}^{(n+1)}$, and then uses the related $a_r^{(n+1)}$ and $b_v^{(n+1)}$ to calculate $g^{(n+1)}$ via

@@

To broach this complications, we again use a greedy algorithm with degree tempering.

@@

Put another way, rather than two simultaneous greedy optimizations over $R+1$ and $V$ symbols (with $(R+1)V$ iterations), a single greedy process over $V+R+1$ symbols is performed, where each symbol is additionally labeled as being in the numerator or denominator.

@@

Both \gmvp{} and \gmvr{} are publicly available on Github through the \texttt{positive} repository (Ref. \cite{lionel_london_2018_1402516}), and may be imported in python via {\small{\texttt{positive.learning.gmvpfit}}} and {\small{\texttt{positive.learning.gmvrfit}}}.

@@

Second, we apply \gmvr{} to the modeling of spin -2 spherical-spheroidal harmonic mixing coefficients (\ceqn{sigma}).

@@

For relevance of presentation to physics examples in subsequent sections, we will restrict ourselves to a case where numerical noise is low, and the negative greedy step does not alter the output of Alg. \ref{alg:gmvr}.

@@

\par To generate the test data, \eqn{eq:gmvrtoy1} is evaluated with 25 points in $x_0$ and $x_1$, where each is between -3 and 3. Though not a requirement of \gmvr{}, for simplicity of presentation, domain points are equally spaced.

@@

\par In this rudimentary example case, \gmvr{} correctly recovers the functional form of the input data, and accurately recovers the correct values of model parameters.

@@

But, in general, \gmvr{} and related techniques, having no knowledge of the underlying noise distribution, will attempt to model minor correlations and offsets within the training data's noise.

@@

However, we have demonstrated the utility of \gmvr{} in a relatively ideal usage case where the underlying function is rational, and the training data is only weakly contaminated with noise.
\par In the following sections, we consider realistic, but similarly ideal cases, where the functional form of the sample data is not known to be explicitly polynomial or rational, but the amount of noise within the training data is negligible.

@@

\subsection{Modeling \qnm{} frequencies with \gmvp{}}

@@

\par \fig{fig:qnm}'s top left and right panels' upper axes demonstrate the effect of mapping $\jf$ onto $\kappa$.

@@

In particular, it is shown that the two branches (namely $\jf>0$ and $\jf<0$) naturally form a single family of solutions when accounting for the sign of the \bh{}'s oriented spin \cite{Husa:2007hp}.

@@

Concurrently, the use of $\kappa$ as a domain variable has the desirable effect of making each $\cw\lmn$ and $\tau\lmn$ approximately polynomial.

@@

\subsection{Modeling spherical-spheroidal inner-products with \gmvr{}}

@@

For efficiency of presentation, each $\sigma_{\LMlmn}$ is plotted via its real and imaginary part.

@@

In particular, the \gmvp{} algorithm performs multivariate polynomial modeling of real and complex valued scalar functions with no inherent limitation on the number of domain parameters.

@@

When applied to the modeling of analytically computed quantities, both algorithms perform extremely well in producing accurate and precise representations of training data, suggesting extended applicability of \gmvr{} and \gmvp{} similar problems.

@@

\par Treating a toy problem with \gmvr{} demonstrates its ability to faithfully recover underlying model parameters for a plausible dataset.

@@

\par Both \gmvp{} and \gmvr{} may be used to automatically determine the functional form and model for a given dataset that is expected to be respectively polynomial or rational.

@@

The fits presented are of direct use in Ref. \cite{London:2017bcn}, where efficiently evaluable \qnm{} frequencies are required to general template waveforms for \gw{} searches and parameter estimation.
The fits presented may find future use in Phenom or EOB based \gw{} models.
\par \gmvr{} has been applied to the modeling of mixing coefficients between the spherical and spheroidal harmonics.
These fits are of direct use in Ref. \cite{Carullo:2018sfu}, and may be of future use in similar ringdown-only models for the purpose of testing \gr{}.
\par While \gmvr{} and \gmvp{} show promise in the cases shown here, in their presented rudimentary form, both posses a number of limitations.
Neither method directly accounts for information
